---
title: "SI Course Project (Part 1)"
author: "Cecilia Cruz-Ram, MD DPCOM"
date: "4/2/2018"
output: pdf_document
---

## Introduction:
In this project we will investigate the exponential distribution in R and compare it with the 
<b>Central Limit Theorem</b>. The exponential distribution can be simulated in R with 
<b>rexp(n, lambda)</b> where lambda is the rate parameter. The mean of exponential distribution 
is 1/lambda and the standard deviation is also 1/lambda. 

Provided Data:
* lambda = 0.2 for all of the simulations
* Distribution of averages of 40 exponentials
* Number of simulations = 1000

## Simulations:

A. Setwd
```{r}
setwd("/Users/sexybaboy/Documents/Files/Zetch/Online Courses/Data Science Specialization Feb18/R/Statistical Inference")
```

B. Set seed of reproducibility
```{r}
set.seed(2018)
```

C. Set sample size
```{r}
n <- 40              # sample size
lambda <- 0.2        # number of exponentials
simNum <- 1000       # number of simulations
```

D. Run simulations
```{r}
simMeans = NULL
for (i in 1 : 1000) {
  simMeans = c(simMeans, mean(rexp(n,lambda)))
}

head(simMeans)
```

E. Instructions

<b>*1. Show the sample mean and compare it to the theoretical mean of the distribution.*</b>

Run theoretical mean
```{r}
theosampMean <- round(1/lambda,3)
theosampMean
```

Run actual mean
```{r}
actualMean <- round(mean(simMeans),3)
actualMean
```

Plot showing both means
```{r}
require(ggplot2)
simMeansDf <- as.data.frame(simMeans)
g <- ggplot(simMeansDf, aes(x = simMeans))
g <- g + geom_histogram(binwidth = .3, color = "black", fill = "white") +
  geom_vline(xintercept = theosampMean, color = "red", size = 1, linetype = 2) +
  geom_vline(xintercept = actualMean, color = "blue", size = 1, linetype = 3) +
  labs(x = "Simulation Means", y = "Frequency", 
  title = "Theoretical vs Actual Mean")
g
```

*The red dashed vertical line indicate the theoretical sample mean, 1/lambda = 5, 
while the green dashed vertical line is the calculated average sample mean size 
of 40 of 1000 samples showing very close proximity.*

<b>*2. Show how variable the sample is (via variance) and compare it to the theoretical 
variance of the distribution.*</b>

Run theoretical variance
```{r}
theosampVar <- round((1/lambda)^2/n,3)
theosampVar
```

Run actual variance
```{r}
actualVar <- round(var(simMeans),3)
actualVar
```

Table showing both theoretical and actual mean and variance
```{r}
tab <- matrix(c(theosampMean, actualMean, 
                theosampVar,actualVar),
              ncol = 2, byrow = TRUE)
colnames(tab) <- c("Theoretical","Sample")
rownames(tab) <- c("Mean","Variance")
tab <- as.table(tab)
tab
```

*Preceding table shows near identical values.*

<b>*3. Show that the distribution is approximately normal.*</b>

<b>In point 3, focus on the difference between the distribution of a large collection 
of random exponentials and the distribution of a large collection of averages of 40 exponentials.*</b>

Make a histogram with the density and sample means. Add density curve of the normal distribution 
and the sample distribution:
```{r}
g <- ggplot(simMeansDf, aes(x=simMeans))
g <- g + geom_histogram(binwidth = .2, color = "black", fill = "white" , aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = theosampMean, sd = sd(simMeans)), 
                color = "red", size = 1) +
  stat_density(geom = "line", color = "blue", size = 1)  +
  labs(x = "Simulation Means", y = "Density", 
       title = "Density of Simulated Exponential Means")
g
```

*Plot2 shows the distribution of means of the sampled exponential distributions which 
appear to follow a normal distribution, due to the <b>Central Limit Theorem</b>. 
An increase in the number of samples (currently 1000) will create a distribution 
that would be even closer to the standard normal distribution. The red line above is 
the normal distribution curve which closely approximates the blue colored sample curve.*