---
title: "SI Course Project (Part 2)"
author: "Cecilia Cruz-Ram, MD DPCOM"
date: "4/7/2018"
output: pdf_document
fontsize: 9pt
geometry: margin=0.5in
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Basic Inferential Data Analysis**
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

#### A. Setwd and load package
```{r}
setwd("/Users/sexybaboy/Documents/Files/Zetch/Online Courses/Data Science Specialization Feb18/R/Statistical Inference")
require(ggplot2)
```
#### B. Instructions
##### 1. Load the ToothGrowth data and perform some basic exploratory data analyses.
Load ToothGrowth data
```{r}
data("ToothGrowth")
```

Show Unique Values
```{r}
unique(ToothGrowth$len)
unique(ToothGrowth$supp)
unique(ToothGrowth$dose)
```
Analyze and explore data via plots
Convert dose to a factor
```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```
Plot tooth length ('len') vs. the dose amount ('dose'), broken out by supplement delivery method ('supp')
```{r}
bp <- ggplot(ToothGrowth, aes(x = dose, y = len, fill = dose)) +
  geom_boxplot(position = "dodge", varwidth = TRUE) + xlab("Dose Amount") + ylab("Tooth Length") + facet_grid(~ dose) + 
  ggtitle("Tooth Length v Dose Amount (by Delivery Method)") + 
  theme(plot.title = element_text(lineheight = .3))
bp + scale_fill_hue(l = 70, c = 50)
```

Plot tooth length ('len') vs. supplement delivery method ('supp') broken out by the dose amount ('dose')
```{r}
bp <- ggplot(ToothGrowth, aes(x = supp, y = len, fill = dose)) +
  geom_boxplot(position = "dodge", varwidth = TRUE) + xlab("Supplement Delivery") + ylab("Tooth Length") + facet_grid(~  supp) + ggtitle("Tooth Length v Delivery Method (by Dose Amount)") + 
  theme(plot.title = element_text(lineheight = .3))
bp + scale_fill_hue(l = 70, c = 50)
```

##### 2. Provide a basic summary of the data.
```{r}
summary(ToothGrowth)
```
##### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
Run t-test
```{r}
t.test(len~supp, data = ToothGrowth)
```
Run t-test using dose amounts 0.5 and 1.0
```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0, 0.5))
t.test(len~dose, data = ToothGrowth_sub)
```
Run t-test using dose amounts 0.5 and 2.0
```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5, 2.0))
t.test(len~dose, data = ToothGrowth_sub)
```
Run t-test using dose amounts 1.0 and 2.0
```{r}
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0, 2.0))
t.test(len~dose, data = ToothGrowth_sub)
```
##### 4. State your conclusions and the assumptions needed for your conclusions.

###### Assuming the sample is representative of the population and the distribution of the sample means follows the Central Limit Theorem, we found out that the higher the dosage, the longer the tooth grows over time. Supplement delivery method, however, has no effect on tooth growth/length.